Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-06 WSSSPE1_report/WSSSPE1_summary_report.tex
more work in sustainability section (8)

@@

In the Developing and Maintaining Software panel,

\begin{itemize}
\item what does sustainable mean?
\begin{itemize}
\item from http://oss-watch.ac.uk/resources/ssmm: When choosing
  software for procurement or development reuse - regardless of the
  license and development model you will use - you need to consider
  the future. While a software product may satisfy today's needs, will
  it satisfy tomorrow's needs? Will the supplier still be around in
  five years' time? Will the supplier still care for all its customers
  in five years' time? Will the supplier be responsive to bug reports
  and feature requests? In other words, is the software sustainable?
\item I think this is a key question we as a community need to agree
  on (whats our problem statement?) and hope we come out today with
  some progress. - @VonWelch
\item What is the goal of sustainability? Reproducible science,
  persistence, quality, or else?
\item How do you measure success? How do you know when you have
  actually achieved sustainable software?
\item Sustainability should be addressed throughout the life cycle
  (Daisie Huang)
\end{itemize}

\item What is the relationship between sustainability and provenance?

\item Is usability separate from sustainability or a fundamental part?
\end{itemize}

>>>>

In the Developing and Maintaining Software panel, there was some discussion
on this topic: what does sustainability mean? It was pointed out that
OSS Watch proposes a Software Sustainability Maturity Model to address
the issue of how sustainability a particular element of software is, and says
that this sustainability is important. ``When choosing
software for procurement or development reuse - regardless of the
license and development model you will use - you need to consider
the future. While a software product may satisfy today's needs, will
it satisfy tomorrow's needs? Will the supplier still be around in
five years' time? Will the supplier still care for all its customers
in five years' time? Will the supplier be responsive to bug reports
and feature requests? In other words, is the software sustainable?''~\cite{OSS-ssmm-web}

Attendees suggested that a key question that the definition of sustainability is
one issue on which the community needs to agree, and that ideally, an initial
definition would be determined during the workshop, or at least some progress
would be made towards this goal.  Another topic that came up is what the
goal of sustainability is.  Perhaps it's reproducible science, or
persistence, or quality, or something else.  Similarly, some attendees want
to understand how success in sustainability is measured.  How does a set of
developers know when they have actually achieved sustainable software?
This led to a comment that sustainability should be addressed throughout the
full software life cycle.

Another topic that came up during the Developing and Maintaining Software
panel is the relationship of sustainability to other software attributes.  Attendees
asked: what is the relationship between sustainability and provenance? And:
is usability separate from sustainability or a fundamental part of it?

@@
